---
title: "Gov 50 Final Project"
author: "Vanessa Iwuoha"
description: "My final project: Exploring the stark Social Mobility discrepancies between with two Newark Wards: East and Central"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction
describe the research question and main hypothesis; describe why it is important.


##Data section 
2-3 paragraphs + plot visualizing main outcome of interest. (3pts)

```{r}
```


##Results section 
plot of main analysis + regression output + 2-3 paragraphs of description and interpretation of the plots and regression (including interpreting the main coefficient of interest and describing if it is statistically significant and if we should interpret it causally). This section could be longer if you choose to include additional analyses. (8pts)

```{r}
```


##Conclusion section
1 paragraph (i) summarizing results and assessing the extent to which you find support for your hypothesis; (ii) describing limitations of the analysis and threats to inference (missing data, confounding, etc), and stating how you could improve your analysis if you had more time/money. (2pts)


#Extra Analysis/ Thinking Process 

```{r}
```
