---
title: "Gov 50 Final Project"
author: "Vanessa Iwuoha"
description: "My final project: Exploring the stark Social Mobility discrepancies between with two Newark Wards: East and Central"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction

```{r}
library(tidyverse)
library(haven)
#ec50 data 
atlas <- read_dta("~/Downloads/atlas.dta") #this has issues loading
#social capital data
social_capital <- read_csv("~/Downloads/social_capital_zip.csv")

#filtering the atlas data
row_numbers <- grepl(pattern = "Newark, NJ", x = atlas$tract_name)
Newark_with_East <- subset(x = atlas, subset = row_numbers)
east_newark_rows <- grepl(pattern = "East Newark, NJ", x = Newark_with_East$tract_name)
newark <- subset(x = Newark_with_East, subset = !east_newark_rows)

#separation by the ward
newark <- newark |>
  mutate(
    wards = case_when(
      tract_name %in% c("Forest Hill, Newark, NJ", "Lower Roseville, Newark, NJ", "Upper Roseville, Newark, NJ", "Mount Pleasant/Lower Broadway, Newark, NJ", "North Broadway, Newark, NJ", "Seventh Avenue, Newark, NJ") ~ "North",
      
      tract_name %in% c("Upper Clinton Hill, Newark, NJ", "Weequahic, Newark, NJ", "Dayton/Weequahic Park, Newark, NJ", "South Broad Street, Newark, NJ") ~ "South",
      
      tract_name %in% c("Fairmount, Newark, NJ", "Lower Vailsburg, Newark, NJ", "Upper Vailsburg, Newark, NJ", "West Side, Newark, NJ") ~ "West",
      
      tract_name %in% c("Ironbound District, Newark, NJ", "North Ironbound, Newark, NJ") ~ "East",
      
      tract_name %in% c("University Heights, Newark, NJ", "Springfield/Belmont, Newark, NJ", "Central Business District, Newark, NJ", "University Heights, Newark, NJ") ~ "Central", 
    ))

#filtering the social capital data
newark_2 <- social_capital |>
  filter(zip %in% c(7102, 7103, 7104, 7105, 7106, 7107, 7108, 7112, 7114))

#separating by ward > through futher digging, i saw that one zipcode could contain many neighborhoods
newark_2 <- newark_2 |>
  mutate(
   wards = case_when( 
      zip %in% c(7104, 7107) ~ "North",
      zip %in% c(7112, 7108, 7114) ~ "South",
      zip %in% c(7105) ~ "East",
      zip %in% c(7106, 7103) ~ "West",
      zip %in% c(7102) ~ "Central",
    ))

#combining the data set 
Newark <- newark |>
  left_join(newark_2, join_by(wards))

```

Last semester, I researched social mobility in my hometown, Newark, New Jersey, using Raj Chetty's data from Opportunity Insights. Through researching different variables across my city's wards, or neighborhoods, I found that two had extremely opposite outcomes. The East Ward, with a majority white and Hispanic population, had the highest median income and social mobility rating. The Central Ward, on the other hand, has a majority Black population and has the social mobility score and second lowest median income.

Those observations led me to my research question: Why is there a stark difference between the social mobility outcomes of two wards in Newark, the Central and East? From my preliminary data, I believe the discrepancy in the amount of foreign-born residents might be the cause of social mobility ratings, perhaps due to differences in educational attainment. I also explore if there is a discrepancy between the two ward's "economic connectedness," (a measure of the connection between low and high income residents in an area) an element that Chetty argues is strongly tied to social mobility. This information is incredibly important to understand because Newark is a low-income, disadvantaged city, with a rich history and great potential. Despite numerous historical attempts at revitalization, many residents are living in poverty and the city is still performing poorly. I think this study is a great first to look into potential solutions by understanding why some populations are doing better than others and what lessons could be incorporated from that analysis.

------------------------------------------------------------------------

## Data 

### Background Data Across Wards

```{r, echo = TRUE}
library(kableExtra)

vars_interest <- Newark |>
  group_by(wards)|>
  filter(wards != "") |>
  summarize_at(
    vars(kfr_pooled_pooled_p25, med_hhinc2016, foreign_share2010, singleparent_share2010, share_black2010, share_hisp2010, share_white2010, frac_coll_plus2010),
    mean)

kable (vars_interest, col.names = c("Wards", "Social Mobility", "Median Household Income 2016", "Foreign Share 2010", "Single Parent Share 2010", "Black Share 2010","Hispanic Share 2010", "White Share 2010", "Fraction with College Degree or More 2010"), digits = 2) |>
  kable_styling(font_size = 10)

```

All the data is collected from Chetty's Opportunity Insights. The chart above displays different variables and outcomes across the wards. From this table, we can see that the East ward has the highest foreign share, highest social mobility score, highest white share, and highest median income. The Central ward has the 2nd lowest median income, third highest black share, and 3rd lowest social mobility (there is some discrepancy between the Central results here compared to EC50, due to the data set combination- to be explained in conclusions).

#### Here are descriptions of the data we will be discussing: 

The "kfr_pooled_pooled_p25" is the social mobility measure of absolute mobility for children born in the lower 25 percentile of income. Which, essence, answers the question: "Where do people in the 25th percentile end up?" It is a percentile ranking from 0 to 100, with 0 being the lowest ranking and 100 the highest. In this paper, I will just refer to it as social mobility.

The "foreign_share2010" is the fraction of immigrants.

The following: "share_black2010," "share_white2010," and "share_hisp2010" are all fractions of those races in a given area. 

The "frac_coll_plus2010" is the fraction of those with a college degree or higher in 2010.


### Social Capital Data Across Wards

```{r, echo = TRUE}
social_capital_ward <- Newark |>
  group_by(wards) |>
  filter(wards != "") |>
  summarize_at(
    vars(ec_zip, nbhd_ec_zip, ec_high_zip, exposure_grp_mem_zip, 
exposure_grp_mem_high_zip, volunteering_rate_zip),
    mean)

kable (social_capital_ward, col.names = c("Wards", "High-type Economic Connectedness Calculated using Neighborhood Friends", "Economic Connectedness", "Economic Connectedness for High-SES Individuals", "Mean Exposure to High-SES Individuals for Low-SES Individuals", "Mean Exposure to high-SES Individuals for High-SES Individuals", "Volunteering Rate"), digits = 2) |>
  kable_styling(font_size = 10)

```

This table shows the different measures for economic connectedness across the wards, this element is noted by Chetty to be highly correlated with social mobility. Here we see that there a quite a bit of NA's, but the East ward has the lowest economic connectedness score and the Central ward having the highest "Exposure to High-SES Individuals for Low-SES Individuals" score. I ended up not futher analysizing these correlations due to the lack of data for the most important variable: "ec_zip".

#### Description of variables 
*The descriptions below are taken directly from the Opportunity Insights' Social Capital II: Determinants of Economic Connectedness data set "Read me" file.*

* Social Capital Data: Determinants of Economic Connectedness
  + "ec_zip": baseline definition of economic connectedness
  + "nbhd_ec_zip": economic connectedness calculated using only within-neighborhood friends
  + "ec_high_zip": economic connectedness for high-SES individuals
  + "exposure_grp_mem_zip": mean exposure to high-SES individuals by ZIP code for low-SES individuals
  + "exposure_grp_mem_high_zip": mean exposure to high-SES individuals by ZIP code for high-SES individuals
  + "volunteering_rate_zip": the percentage of Facebook users who are members of a group which is predicted to be about ‘volunteering’ or ‘activism’ based on group title and other group characteristics. 

### Share of College Graduates Across Wards (2010)

```{r, echo = TRUE}
Newark |>
   filter(wards != "") |>
  ggplot(mapping = aes(x = wards, y = frac_coll_plus2010)) +
           geom_col() +
  labs(
    x = "Wards",
    y = "Fraction of College Degree or Higher 2010"
  )

```

The bar plot shows the varying shares of college graduates across the wards with the East having the lowest share, the South having the highest share, and the Central ward having the 2nd lowest share.

------------------------------------------------------------------------

## Results 

### Correlation Between Foreign Share and Social Mobility (2010)

```{r, echo = TRUE}
Newark |>
  ggplot(mapping = aes(x = foreign_share2010, y = kfr_pooled_pooled_p25)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Foreign Share 2010",
    y = "Social Mobility"
  ) 

cor(Newark$foreign_share2010, Newark$kfr_pooled_pooled_p25, use = "complete.obs")

```

This chart shows there is a strong correlation between foreign share and social mobility in Newark: 0.82. It is interesting to think about considering that the East Ward has the highest foreign share as well as the highest social mobility score. However, this is not enough information to establish causation.

### Correlation Between Fraction of Population with Amount of Black Residents and Social Mobility (2010)

```{r}
Newark |>
  ggplot(mapping = aes(x = share_black2010 , y = kfr_pooled_pooled_p25 )) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Fraction of Black Residents 2010",
    y = "Social Mobility"
  ) 

cor(Newark$share_black2010, Newark$kfr_pooled_pooled_p25, use = "complete.obs")
```

There is a strong negative correlation between share black and social mobility: -0.77. This is interesting to note because the wards with higher amounts of black residents, like the Central and South ward have lower social mobility. This, however, is not proof of causation and there most likely is come kind of confounder with the race aspect.

### Correlation Between Fraction of Population with College Degree and Social Mobility (2010)

```{r}
Newark |>
  ggplot(mapping = aes(x = frac_coll_plus2010 , y = kfr_pooled_pooled_p25 )) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Fraction of College Degree or Higher 2010",
    y = "Social Mobility"
  ) 

cor(Newark$frac_coll_plus2010, Newark$kfr_pooled_pooled_p25, use = "complete.obs")

```

Here, we see interestingly enough, the amount of college attainment is not at all correlated with social mobility (0.003). It is near zero, which seems a bit off and might be impacted by some of the missing data. 

### Regression Output 

```{r}
library(broom)
  lm_1 <-lm(kfr_pooled_pooled_p25 ~ foreign_share2010 + share_black2010 + frac_coll_plus2010, data = Newark)
  
modelsummary::modelsummary(lm_1, statistic = c("s.e. = {std.error}", "p = {p.value}"), gof_map = c("nobs", "r.squared", "adj.r.squared"))
  
```

This regression reveals a lot of interesting information. The intercept shows, that holding those 3 variables constant, the social mobility for children in the 25th income percentile in Newark is 33, which is extremely low compared to the country average of *42.85* and the New Jersey average of *46.47* *(data from my EC50 project)* We observe that each 1 increase of foreign share adds an average of 22.52 to that social mobility outcome, holding the other variables constant, and it is statistically significant. We that see that black share reduces social mobility by 6 by each percent increase, controlling for the other variables. The effect is also statistically significant, with both measures having a a p-value of less than 0.001. The fraction of those with a college degree has a mere 4.57 measure effect with each percent increase but it is not statistically significant.

------------------------------------------------------------------------

## Conclusion

Summarize Results 

The plots and linear regression reveals that there is indeed a high correlation between black share and foreign share on social mobility. We also observe that these effects have a high statistical significance. Concerning my hypothesis, I was right to sense a correlation between foreign share and social mobility, but I was interestingly enough, unable to like it to educational attainment. There was a low correlation between fraction of college degrees and social mobility. I was also unable to test the economic connectedness hypothesis due to scarce data. Some of the limitations of my study are that the combination of the data sets led to a mismatch between how the different wards were created. In the “atlas” data, I created the wards based on the census tracts because it included the specific neighborhoods, but in the “social capital” data, the wards were created based on zip codes, which sometimes overlaps between wards. This also caused some data like social mobility to very form what I found in my original EC50 project. I solved this by grouping the wards of the “social capital” data based on which zip codes had the largest amount of neighborhoods in a certain ward. A huge issue is that “foreign share” and “black share,” though they are highly statistically significant, could be confounders. There could be other factors such as education or occupation, that may be the real “cause” of social mobility but are correlated with variables like race. If I had more time and resources, I would collect more detailed/individualized census data. I would gather data on specified immigrant groups (specific to region in the world and maybe even by country), SAT scores, parent occupation and college attainment, and the type of high schools attended by the children (charter school, magnet school, or regular public school). It would be interesting to pair the data analysis sociology concepts in order to provide further understanding. A big next step would be executing an experiment to test for causation between "foreign share" and social mobility.
