---
title: "Gov 50 Final Project"
author: "Vanessa Iwuoha"
description: "My final project: Exploring the stark Social Mobility discrepancies between with two Newark Wards: East and Central"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction

```{r, echo = FALSE}
library(tidyverse)
library(haven)
#ec50 data 
atlas <- read_dta("~/Downloads/atlas.dta") #this has issues loading
#social capital data
social_capital <- read_csv("~/Downloads/social_capital_zip.csv")

#filtering the atlas data 
row_numbers <- grepl(pattern = "Newark, NJ", x = atlas$tract_name)
Newark_with_East <- subset(x = atlas, subset = row_numbers)
east_newark_rows <- grepl(pattern = "East Newark, NJ", x = Newark_with_East$tract_name)
newark <- subset(x = Newark_with_East, subset = !east_newark_rows)

#separation by the ward
newark <- newark |>
  mutate(
    wards = case_when(
      tract_name %in% c("Forest Hill, Newark, NJ", "Lower Roseville, Newark, NJ", "Upper Roseville, Newark, NJ", "Mount Pleasant/Lower Broadway, Newark, NJ", "North Broadway, Newark, NJ", "Seventh Avenue, Newark, NJ") ~ "North",
      
      tract_name %in% c("Upper Clinton Hill, Newark, NJ", "Weequahic, Newark, NJ", "Dayton/Weequahic Park, Newark, NJ", "South Broad Street, Newark, NJ") ~ "South",
      
      tract_name %in% c("Fairmount, Newark, NJ", "Lower Vailsburg, Newark, NJ", "Upper Vailsburg, Newark, NJ", "West Side, Newark, NJ") ~ "West",
      
      tract_name %in% c("Ironbound District, Newark, NJ", "North Ironbound, Newark, NJ") ~ "East",
      
      tract_name %in% c("University Heights, Newark, NJ", "Springfield/Belmont, Newark, NJ", "Central Business District, Newark, NJ", "University Heights, Newark, NJ") ~ "Central", 
    ))

#Filter the data by wards (atlas data)
#filtering the social capital data
newark_2 <- social_capital |>
  filter(zip %in% c(7102, 7103, 7104, 7105, 7106, 7107, 7108, 7112, 7114))

#separating by ward > through futher digging, i saw that one zipcode could contain many neighborhoods
newark_2 <- newark_2 |>
  mutate(
   wards = case_when( 
      zip %in% c(7104, 7107) ~ "North",
      zip %in% c(7112, 7108, 7114) ~ "South",
      zip %in% c(7105) ~ "East",
      zip %in% c(7106, 7103) ~ "West",
      zip %in% c(7102) ~ "Central",
    ))

#combine 
Newark <- newark |>
  left_join(newark_2, join_by(wards))

```

Last semester, I researched social mobility in my hometown, Newark, New Jersey, using Raj Chetty's data from Opportunity Insights. Through researching different variables amongst my city's wards, or neighborhoods, I found that two had extremely opposite outcomes. The East Ward, with a majority white and Hispanic population, had the highest median income and social mobility rating. The Central Ward, on the other hand, has a majority Black population and has the social mobility score and second lowest median income.

Those observations led me to my research question: Why is there a stark difference between the social mobility outcomes of two wards in Newark, the Central and East? From my preliminary data, I believe the discrepancy in the amount of foreign-born residents might be the cause of social mobility ratings, perhaps due to differences in educational attainment. I also explore if there is a discrepancy between the two ward's "economic connectedness," (a measure of the connection between low and high income residents in an area) an element that Chetty argues is strongly tied to social mobility. This information is incredibly important to understand because Newark is a low-income, disadvantaged city, with a rich history and great potential. Despite numerous historical attempts at revitalization, many residents are living in poverty and the city is still performing poorly. I think this study is a great first to look into potential solutions by understanding why some populations are doing better than others and what lessons could be incorporated from that analysis.

------------------------------------------------------------------------

## Data 

### Background Data Across Wards

```{r, echo = FALSE}
library(kableExtra)

#Background to compare variables and outcomes across 
vars_interest <- Newark |>
  group_by(wards)|>
  filter(wards != "") |>
  summarize_at(
    vars(kfr_pooled_pooled_p25, med_hhinc2016, foreign_share2010, singleparent_share2010, share_black2010, share_hisp2010, share_white2010, frac_coll_plus2010),
    mean)

kable (vars_interest, col.names = c("Wards", "Social Mobility", "Median Household Income 2016", "Foreign Share 2010", "Single Parent Share 2010", "Black Share 2010","Hispanic Share 2010", "White Share 2010", "Fraction with College Degree or More 2010"), digits = 2) |>
  kable_styling(font_size = 10)

```

All the data is collected from Chetty's Opportunity Insights. The chart above displays different variables and outcomes across the wards. From this table, we can see that the East ward has the highest foreign share, highest social mobility score, highest white share, and highest median income. The Central ward has the 2nd lowest median income, third highest black share, and 3rd lowest social mobility (there is some discrepancy between the central results here compared to EC50, due to the data set - to be explained in conclusions).


### Social Capital Data Across Wards

```{r, echo = FALSE}
social_capital_ward <- Newark |>
  group_by(wards) |>
  filter(wards != "") |>
  summarize_at(
    vars(ec_zip, nbhd_ec_zip, ec_high_zip, exposure_grp_mem_zip, 
exposure_grp_mem_high_zip, volunteering_rate_zip),
    mean)

kable (social_capital_ward, col.names = c("Wards", "High-type Economic Connectedness Calculated using Neighborhood Friends", "Economic Connectedness", "Economic Connectedness for High-SES Individuals", "Mean Exposure to High-SES Individuals for Low-SES Individuals", "Mean Exposure to high-SES Individuals for High-SES Individuals", "Volunteering Rate"), digits = 2) |>
  kable_styling(font_size = 10)

```

This data shows the different measures for economic connectedness across the wards, this element is noted by Chetty to be highly correlated with social mobility. Here we see that there a quite a bit of NA's, but the East ward has the lowest economic connectedness score and the Central ward having the highest "Exposure to High-SES Individuals for Low-SES Individuals" score.

### Share of College Graduates Across Wards (2010)

```{r, echo = FALSE}
#college education by ward - trying to figure out
Newark |>
   filter(wards != "") |>
  ggplot(mapping = aes(x = wards, y = frac_coll_plus2010)) +
           geom_col() +
  labs(
    x = "Wards",
    y = "Fraction of College Degree or Higher 2010"
  )
```

The bar plot shows the varying shares of college graduates across the wards with the East having the lowest share, the South having the highest share, and the Central ward having the 2nd lowest share.

------------------------------------------------------------------------

## Results 

plot of main analysis + regression output + 2-3 paragraphs of description and interpretation of the plots and regression (including interpreting the main coefficient of interest and describing if it is statistically significant and if we should interpret it causally). This section could be longer if you choose to include additional analyses. (8pts)

### Correlation Between Foreign Share and Social Mobility (2010)

```{r, echo = FALSE}
#plot of main analysis
#plot of interest: there is a high correlation between foreign share and social mobility

Newark |>
  ggplot(mapping = aes(x = foreign_share2010, y = kfr_pooled_pooled_p25)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Foreign Share 2010",
    y = "Social Mobility"
  ) 

#how to show correlation number on the chart??
cor(Newark$foreign_share2010, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #0.8285084

```

This chart shows there is a strong correlation between foreign share and social mobility in Newark. This is interesting to think about considering that the East Ward has the highest foreign share as well as the highest social mobility scor, which we know is highly correlated.

### Correlation Between Fraction of Population with College Degree and Social Mobility (2010)

```{r, echo = FALSE}
#no correlation between frac college and social mobility 
Newark |>
  ggplot(mapping = aes(x = frac_coll_plus2010 , y = kfr_pooled_pooled_p25 )) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Fraction of College Degree or Higher 2010",
    y = "Social Mobility"
  ) 

```

### Regression Output 

```{r, echo = FALSE}
#regression output
library(broom)
  lm_1 <-lm(kfr_pooled_pooled_p25 ~ foreign_share2010 + share_black2010 + frac_coll_plus2010, data = Newark)
  
  tidy(lm_1)|>
    knitr::kable(digits = 3)
  
#how should i interpret the wards section
  
```

------------------------------------------------------------------------

## Conclusion 

1 paragraph (i) summarizing results and assessing the extent to which you find support for your hypothesis;

RESULTS -> its more about 
black and foreign share might be covriates for something else

(ii) describing limitations of the analysis and threats to inference (missing data, confounding, etc),

(iii) and stating how you could improve your analysis if you had more time/money.
