---
title: "Gov 50 Final Project"
author: "Vanessa Iwuoha"
description: "My final project: Newark Data"
output:
  distill::distill_article:
    self_contained: false
---



## Milestone 1: Create public GitHub repository

I am interested in exploring data related to something related to education or social mobility in my city, Newark

making a change to try and update html

## Milestone 2: Finding data and writing a proposal 

1. What is the cause of starting differences in racial makeup and social mobility across different Newark wards?
2. The cause of huge differences in racial population across different Newark wards may be a product of different immigrant groups.
3. My variables of interests would be the different immigrant populations (by country/continent/region) in different zip codes/neighborhoods/wards in Newark. I would also look into the “Absolute Mobility at the 25th Percentile” measure by Raj Chetty.
4. My outcome variable would be the “Absolute Mobility at the 25th Percentile” measure by Raj Chetty. It is the average rank of the children whose parents are at the 25 percentile in the income ranking. It is measured on a scale from 0 - 100.
5. To support my hypothesis, I would have to see a high correlation between immigration and absolute mobility. A negative or positive correlation would detail whether certain populations have a “positive” or “negative effect” on absolute mobility. My hypothesis would be disproved if there was no correlation.



```{r}
library(tidyverse)
#load raj chetty data 
library(readr)
tract_covariates <- read_csv("~/Library/Mobile Documents/com~apple~CloudDocs/Downloads/tract_covariates.csv")

tract_outcomes_simple <- read_csv("~/Library/Mobile Documents/com~apple~CloudDocs/Downloads/tract_outcomes_simple.csv")
```

```{r}
#combine the data 
data <- tract_covariates |>
  inner_join(tract_outcomes_simple) 
```

```{r}
data_nj <- tract_covariates |>
  inner_join(tract_outcomes_simple) |>
  filter(state == 34) 
  
data_nj |>
  count(czname) #why is it just those 3?? Need to relook at this data 


```


```{r}
newark <- data_nj |>
  filter(county == 13) |> #filter to just nj & essex county
  filter(czname == "Newark") #filter to newark
```


## Milestone 3: One data visualization

The next milestone will require that your Distill article loads the data you have selected and produces one interesting and polished data visualization. This could either show the distribution of one variable or the relationship between two variables.

```{r}
#plot mean household income in 2000 across nj

social_mobility_nj <- data_nj |>
  group_by(county) |>
  mutate(mean_hinc_mean2000 = mean(hhinc_mean2000)) |>
  ggplot(mapping = aes(x = county, y = mean_hinc_mean2000)) +
  geom_point() +
  labs(
    x = "County Number",
    y = "Mean Household Income (2000)"
  )

social_mobility_nj
```

```{r}

library(scales)

#mean household income in 2000 vs single parent share in newark

plot <- newark |>
  ggplot(
    mapping = aes(x = singleparent_share2000, y = hhinc_mean2000)) +
  geom_point() +
  labs(
    x = "Single Parent Share (2000)",
    y = "Mean Household Income (2000)"
    ) +
  scale_y_continuous(labels = comma)

plot #these numbers are definitely WAY TOO BIG, not sure why
```


## Milestone 4: Add results from one analysis 

By this time, your article should contain one visualization and one analysis that attempts to answer your research question. There does not need to be a long discussion, but the results should be presented in either a second visualization or a nicely formatted table.


```{r}
#mean household income in 2000 vs single parent share in newark

plot <- newark |>
  ggplot(
    mapping = aes(x = singleparent_share2000, y = hhinc_mean2000)) +
  geom_point() +
  labs(
    x = "Single Parent Share (2000)",
    y = "Mean Household Income (2000)"
    ) +
  scale_y_continuous(labels = comma)
```

Analysis: There is a clear link between single parent share and mean household income. 

