---
title: "New Project Idea"
author: "Vanessa Iwuoha"
date: "2023-12-06"
output: html_document
---

# Adding the data 

```{r}
library(tidyverse)
library(haven)
#load raj chetty data

#ec50 data 
atlas <- read_dta("~/Downloads/atlas.dta") #this has issues loading

#social capital data
social_capital <- read_csv("~/Downloads/social_capital_zip.csv")

```


#Filter the data by wards (atlas data)
Make the wards separately then combine !
```{r}
#filtering the atlas data 
row_numbers <- grepl(pattern = "Newark, NJ", x = atlas$tract_name)
Newark_with_East <- subset(x = atlas, subset = row_numbers)
east_newark_rows <- grepl(pattern = "East Newark, NJ", x = Newark_with_East$tract_name)
newark <- subset(x = Newark_with_East, subset = !east_newark_rows)

#separation by the ward
newark <- newark |>
  mutate(
    wards = case_when(
      tract_name %in% c("Forest Hill, Newark, NJ", "Lower Roseville, Newark, NJ", "Upper Roseville, Newark, NJ", "Mount Pleasant/Lower Broadway, Newark, NJ", "North Broadway, Newark, NJ", "Seventh Avenue, Newark, NJ") ~ "North",
      
      tract_name %in% c("Upper Clinton Hill, Newark, NJ", "Weequahic, Newark, NJ", "Dayton/Weequahic Park, Newark, NJ", "South Broad Street, Newark, NJ") ~ "South",
      
      tract_name %in% c("Fairmount, Newark, NJ", "Lower Vailsburg, Newark, NJ", "Upper Vailsburg, Newark, NJ", "West Side, Newark, NJ") ~ "West",
      
      tract_name %in% c("Ironbound District, Newark, NJ", "North Ironbound, Newark, NJ") ~ "East",
      
      tract_name %in% c("University Heights, Newark, NJ", "Springfield/Belmont, Newark, NJ", "Central Business District, Newark, NJ", "University Heights, Newark, NJ") ~ "Central", 
    ))

```


#Filter the data by wards (atlas data)

```{r}
#filtering the social capital data
newark_2 <- social_capital |>
  filter(zip %in% c(7102, 7103, 7104, 7105, 7106, 7107, 7108, 7112, 7114))

#separating by ward > through futher digging, i saw that one zipcode could contain many neighborhoods
newark_2 <- newark_2 |>
  mutate(
   wards = case_when( 
      zip %in% c(7104, 7107) ~ "North",
      zip %in% c(7112, 7108, 7114) ~ "South",
      zip %in% c(7105) ~ "East",
      zip %in% c(7106, 7103) ~ "West",
      zip %in% c(7102) ~ "Central",
    ))
```

#Combining the data 

```{r}
#HELP - Did the combination work
Newark <- newark |>
  left_join(newark_2, join_by(wards))
```

#Look into variables to test (grouped by ward)

```{r}
first_check <- Newark |>
  group_by(wards)|>
  summarize_at(
    vars(kfr_pooled_pooled_p25, med_hhinc1990, med_hhinc2016, hhinc_mean2000, frac_coll_plus2000, frac_coll_plus2010, foreign_share2010, share_black2010, share_hisp2010, share_white2010, gsmn_math_g3_2013, singleparent_share1990, singleparent_share2000, singleparent_share2010, emp2000, popdensity2000, popdensity2010, jail_pooled_pooled_p25 ),
    mean)

vars_interest <- Newark |>
  group_by(wards)|>
  summarize_at(
    vars(kfr_pooled_pooled_p25, med_hhinc2016, hhinc_mean2000, frac_coll_plus2010, foreign_share2010, share_black2010, share_hisp2010, share_white2010, gsmn_math_g3_2013, singleparent_share2010, emp2000,jail_pooled_pooled_p25 ),
    mean)
vars_interest


social_capital_ward <- Newark |>
  group_by(wards) |>
  summarize_at(
    vars(ec_zip, nbhd_ec_zip, ec_high_zip, exposure_grp_mem_zip, 
exposure_grp_mem_high_zip, volunteering_rate_zip),
    mean)


#correlations <- Newark |>
  #group_by(wards) |>
  #cor(foreign_share2010, kfr_pooled_pooled_p25, use = "complete.obs")
  
 cor(Newark$foreign_share2010, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #0.8285084
 cor(Newark$foreign_share2010, Newark$singleparent_share2010, use = "complete.obs") #-0.5843314
 cor(Newark$foreign_share2010, Newark$ec_zip, use = "complete.obs") #-0.2888062
 cor(Newark$foreign_share2010, Newark$jail_pooled_pooled_p25, use = "complete.obs") #-0.5725961
  cor(Newark$foreign_share2010, Newark$emp2000, use = "complete.obs") #0.3465099
  cor(Newark$foreign_share2010, Newark$poor_share2010, use = "complete.obs") # -0.3796476
 
 cor(Newark$singleparent_share2010, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.6393986
 cor(Newark$ec_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.359715
 cor(Newark$volunteering_rate_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.3386356
 cor(Newark$exposure_grp_mem_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.5087451
 cor(Newark$nbhd_ec_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.3311841
 cor(Newark$clustering_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #0.4364739
 cor(Newark$frac_coll_plus2010, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #0.003872987
cor(Newark$ec_high_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") # -0.1893304
cor(Newark$share_black2010, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.7738769

 
 #not sure why it doesnt work lol
 cor(Newark$fgsmn_math_g3_2013, Newark$kfr_pooled_pooled_p25, use = "complete.obs")
 cor(Newark$singleparent_share2010, Newark$fgsmn_math_g3_2013, use = "complete.obs")
 
 Newark|>
   count(gsmn_math_g3_2013)
 
 Newark|>
   mean(gsmn_math_g3_2013) #why is it NA??
 
 #look at race
 cor(Newark$share_black2010, Newark$poor_share2010, use = "complete.obs") #0.1434808
 cor(Newark$share_black2010, Newark$singleparent_share2010, use = "complete.obs") #0.4707044
 cor(Newark$share_black2010, Newark$jail_pooled_pooled_p25 , use = "complete.obs") #  0.6430772
 
 cor(Newark$share_white2010, Newark$poor_share2010, use = "complete.obs") #-0.2453565
 cor(Newark$share_white2010, Newark$singleparent_share2010, use = "complete.obs") #-0.5488667
 cor(Newark$share_white2010, Newark$jail_pooled_pooled_p25, use = "complete.obs") # -0.510052
 
 cor(Newark$share_hisp2010, Newark$poor_share2010, use = "complete.obs") #-0.05337429
 cor(Newark$share_hisp2010, Newark$singleparent_share2010, use = "complete.obs") #-0.3158382
 cor(Newark$share_hisp2010, Newark$jail_pooled_pooled_p25, use = "complete.obs") # - -0.5502905
 
 #redlining rating 
 redline_score<- Newark |>
  group_by(wards) |> 
   summarize_at(vars(HOLC_A, HOLC_B, HOLC_C, HOLC_D), mean) #na for most
 
 
```

#Try a linear regression 

```{r}

library(broom)
  lm_1 <-lm(kfr_pooled_pooled_p25 ~ foreign_share2010 + singleparent_share2010 + poor_share2010 + jail_pooled_pooled_p25 , data = Newark )
  
  tidy(lm_1)|>
    knitr::kable(digits = 3)
  
```


#bootstrap experiment 
correlation between foreign share and social mobility 
correlation between jail and blackshare

```{r}
```

#final important visuals & stats for hw

```{r}

#Background to compare variables and outcomes across 
vars_interest <- Newark |>
  group_by(wards)|>
  filter(wards != "") |>
  summarize_at(
    vars(kfr_pooled_pooled_p25, jail_pooled_pooled_p25, med_hhinc2016, foreign_share2010, singleparent_share2010, share_black2010, share_hisp2010, share_white2010, frac_coll_plus2010),
    mean)
vars_interest <- vars_interest #HOW TO DELETE LAST ROW

knitr::kable (vars_interest, col.names = c("Wards", "kfr pooled 25", "jail pooled 25", "Median Household Income 2016", "Foreign Share 2010", "Single Parent Share 2010", "Black Share 2010","Hispanic Share 2010", "White Share 2010", "Fraction with College Degree or More 2010"), digits = 2)

#high correlation pairs
 cor(Newark$foreign_share2010, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #0.8285084 GRAPH THIS 
 cor(Newark$foreign_share2010, Newark$med_hhinc2016, use = "complete.obs") #0.443451
 cor(Newark$foreign_share2010, Newark$singleparent_share2010, use = "complete.obs") #-0.5843314
  cor(Newark$foreign_share2010, Newark$jail_pooled_pooled_p25, use = "complete.obs") #-0.5725961
  cor(Newark$foreign_share2010, Newark$poor_share2010, use = "complete.obs") # -0.3796476
 cor(Newark$singleparent_share2010, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.6393986

#Social capital stuff 
social_capital_ward <- Newark |>
  group_by(wards) |>
  filter(wards != "") |>
  summarize_at(
    vars(ec_zip, nbhd_ec_zip, ec_high_zip, exposure_grp_mem_zip, 
exposure_grp_mem_high_zip, volunteering_rate_zip),
    mean)

knitr::kable (social_capital_ward, col.names = c("Wards", "High-type Economic Connectedness Calculated using Neighborhood Friends", "Economic Connectedness", "Economic Connectedness for High-SES Individuals", "Mean Exposure to High-SES Individuals for Low-SES Individuals", "Mean Exposure to high-SES Individuals for High-SES Individuals", "Volunteering Rate"), digits = 2)



#check for if social capital measures positively correlate with social mobility
  cor(Newark$ec_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.359715
  cor(Newark$volunteering_rate_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.3386356
  cor(Newark$exposure_grp_mem_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.5087451
  cor(Newark$nbhd_ec_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #-0.3311841
  cor(Newark$clustering_zip, Newark$kfr_pooled_pooled_p25, use = "complete.obs") #0.4364739
  
#there seems to be negative correlation between "Mean exposure to high-SES individuals for low-SES individuals" and social mobility

  
  
#GRAPHS OF INTEREST
foreign_on_kfr <- Newark |>
  ggplot(mapping = aes(x = foreign_share2010, y = kfr_pooled_pooled_p25)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Foreign Share 2010",
    y = "KFR Pooled 25"
  ) 

foreign_on_jail <- Newark |>
  ggplot(mapping = aes(x = foreign_share2010, y = jail_pooled_pooled_p25)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Foreign Share 2010",
    y = "Jail Pooled 25"
  )

foreign_on_single <- Newark |>
  ggplot(mapping = aes(x = foreign_share2010, y = singleparent_share2010)) + 
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Foreign Share 2010",
    y = "Single Parent Share 2010"
  )


## on education values


#on sytematic racism
   cor(Newark$share_black2010, Newark$singleparent_share2010, use = "complete.obs") #0.4707044
   cor(Newark$share_black2010, Newark$jail_pooled_pooled_p25 , use = "complete.obs") #  0.6430772
   
cor(Newark$share_white2010, Newark$singleparent_share2010, use = "complete.obs") #-0.5488667
 cor(Newark$share_white2010, Newark$jail_pooled_pooled_p25, use = "complete.obs") # -0.510052
 
 cor(Newark$share_hisp2010, Newark$singleparent_share2010, use = "complete.obs") #-0.3158382
 cor(Newark$share_hisp2010, Newark$jail_pooled_pooled_p25, use = "complete.obs") # - -0.5502905
 
```

